Supplementary figures\nCapturing regeneration in models of forest dynamics

### 2023-01-27 manuscript/supplementaryRW.Rmd
adds Landis comment in the report about adding 0 values
added LPJ-GUESS report in the supplementary

---
title: "Supplementary figures\nCapturing regeneration in models of forest dynamics"
author: 
- Olalla Díaz-Yáñez, 
- Yannek Käber, 
- Mats Mahnken, 
- Samuel Fischer, 
- Xavier Morin, 
- Manfred Lexer, 
- Christian Hochauer, 
- Katarina Merganicova, 
- Tobias Mette, 
- Werner Rammer, 
- Kristin H. Braziunas, 
- Paola Mairota, 
- Josef Bruna,
- Heike Lischke, 
- Daniel Scherrer, 
- Simon Scheiter, 
- Thomas Hickler, 
- Tim Anders,
- Jessica Hetzer, 
- Harald Bugmann
date: "`r Sys.Date()`"
output: 
  pdf_document:
    fig_caption: yes
    keep_tex: yes
    number_sections: yes
    latex_engine: xelatex
    extra_dependencies: ["flafter", "float"]
    toc: yes
bibliography: Regeneration.bib
link-citations: yes
linkcolor: blue
csl: global-change-biology.csl
editor_options: 
  markdown: 
    wrap: 72
header-includes:
- \usepackage[left]{lineno}
- \linenumbers
- \newcommand{\beginsupplement}{\setcounter{table}{0}  \renewcommand{\thetable}{S\arabic{table}} \setcounter{figure}{0} \renewcommand{\thefigure}{S\arabic{figure}}}

---

@@

## 4C
### Authors

@@

Morin X., Damestoy T., Toïgo M., Jactel H., Castagneyrol B. Meredieu C. 2020. Using forest gap models and experimental data to explore long-term effects of tree diversity on the productivity of mixed planted forests. Annals of Forest Science. 77 https://doi.org/10.1007/s13595-020-00954-0.

@@

Soil quality is represented by a continuous value between 1 and 5, as provided (see 1.2.3) Soil Field Capacity (SFC) was assessed as follows:
SFC = 8+(X-1)*(25-8)/(5-1), in which X is the ‘soil_qual’ value in the provided ‘soil_quality_200.csv’ file.

@@

Outputs consisted in species specific regeneration rates per decade and per ha, by considering the last 2000 years of simulations to ensure that pseudo-equilibrium state is reached. Results were aggregated for every 10 years for the last 2000 years at the 1ha-level. Variables:
-	Site: numbers corresponding to the climate data file
-	Sample: one per decade, for a total of 2000 years = identified from 1 to 200
-	Species: name specified as requested (eg. ‘fasy’)
-	r.trees: number of trees outcrossing the threshold (7 or 10 cm) during the decade, for the 10 patches simulated (= 1ha), per species
-	r.ba: summed basal area of the r.trees, per species
-	ba: summed basal area of all trees per species at the end of the decade
-	dbh: threshold used

@@

<!--
@article{fischer_lessons_2016,
	title = {Lessons learned from applying a forest gap model to understand ecosystem and carbon dynamics of complex tropical forests},
	volume = {326},
	issn = {03043800},
	url = {https://linkinghub.elsevier.com/retrieve/pii/S0304380015005505},
	doi = {10.1016/j.ecolmodel.2015.11.018},
	language = {en},
	urldate = {2022-09-15},
	journal = {Ecological Modelling},
	author = {Fischer, Rico and Bohn, Friedrich and Dantas de Paula, Mateus and Dislich, Claudia and Groeneveld, J{\"u}rgen and Guti{\'e}rrez, Alvaro G. and Kazmierczak, Martin and Knapp, Nikolai and Lehmann, Sebastian and Paulick, Sebastian and P{\"u}tz, Sandro and R{\"o}dig, Edna and Taubert, Franziska and K{\"o}hler, Peter and Huth, Andreas},
	month = apr,
	year = {2016},
	pages = {124--133},
	file = {Volltext:C\:\\Users\\fischsam\\Zotero\\storage\\EISNTZF6\\Fischer et al. - 2016 - Lessons learned from applying a forest gap model t.pdf:application/pdf},
}
-->

A full description of the model can be found in @fischer_lessons_2016. 
The basis for the parameterization we applied is described in @bohn2014.

@@

To incorporate climate, we used the data set `daily_era5_chelsa_200.csv',
which contains climate data on a daily resolution. We considered the
covariates precipitation, temperature, and irradiance, the latter
of which we converted to units used in Formind. Adding on to these
data, we used a constant and site-independent potential evapotranspiration
(PET) value of $4\frac{\text{mm}}{\text{d}}$ and assumed that the mean day
length was $12\text{h}$ at all sites.

@@

Bugmann, H. (1994). On the ecology of mountainous forests in a changing climate: A simulation study.
Huber, N., Bugmann, H., & Lafond, V. (2020). Capturing ecological processes in dynamic forest models: Why there is no silver bullet to cope with complexity. Ecosphere, 11(5). https://doi.org/10.1002/ecs2.3109

@@

Soil data were translated to bucket sizes values based on the assumpation that the lowest soil quality value of 1 refers to a bucket size of 10 cm and the highest soil quality of 5 refers to a bucket size of 35 cm. Specifically we used the following formula to calculate the bucket size bucket_size = 10 + (soil_quality-1) /4  * 25.

@@

*Please describe how have you calculated each of the variables in the output data.*

@@

@@

### Authors
Heike Lischke

@@

As climatic input, the monthly time series of mean temperature and precipitation sum were used.  With these climatic variables and additional data about slope, aspect and soil field capacity, yearly time series of the following bioclimatic variables were calculated:  (1) DDSum, the yearly day degree sum above 5.5°C , (2) MinWiT, the lowest mean monthly temperature, (3) DrStr, drought stress according to a bucket model  that estimates a monthly PET following (Thornthwaite and Mather 1957), using temperature, latitude, slope and aspect as input,  and calculates the monthly soil water content  based on soil water content in the previous month, precipitation, interception, and field capacity (“bucketsize”)   (Bugmann 1994, Fischlin et al. 1995). Drought stress is then given by 1- the ratio between demand (PET-interception) and the supply, a function of the soil water content.
To get climate data also for the spin-up, bioclimate was calculated from the monthly data   as given, and then extended by sampling from the given data from the end of the climate data to get 1000 years. The same sequence of sampling years was used in all simulations and sites.

@@

The soil bucketsize (available water capacity of the upmost 1m soil layer) required for the drought stress calculation, was derived from the soil quality data set, by
Bucketsize = 10 + (soil_quality-1) /4  * 25 , to covert the range from 10 cm to 35cm

@@

Slope and aspect were taken from the topographic data set and converted to a common variable entering the PET calculation, by
slasp= 2*cos(aspect *pi/180) *min(1, slope/60)

@@

TreeMig simulates the full feedback from seed production to ingrowth. Some of the involved processes depend on environment.
(1) Overall seed production in TreeMig is described by multiplying a species specific seed production per tree that depends on tree size and thus indirectly on the growth, which depends on DDsum, drought stress, nutrients and light and the adult tree number that depends on survival  that depends on DDsum, drought stress, nutrients and light.
(2) Seed dispersal is independent of environment. 
(3) Seed bank dynamics – also environment independent -   is given by seed input, seed mortality, and germination.  Additionally, the number of seeds S of a species in the seed bank of this species (SB) is limited to a carrying capacity C of 1000, by           
SB = C + (SB - C) * Exp(-S/C).  
(4) Germination and seedling survival depend on browsing, drought, winter temperature, DDSum and light. 
(5) The saplings up to 1.37 die and grow like adult trees, depending on DDsum, drought stress and light. Only the parameter value of the light dependence differs from that of the adult trees.  
The regeneration processes were simulated annually in the order 1. germination, 2. germinated seeds leave the seed bank, 3. mortality of seeds in the seed bank, 4. seed entry by seed dispersal. The ingrowth numbers were very sensitive to this order. Simulations with another sequence, in which the seeds in the seed bank died first and then  the  remaining ones germinated,  resulted in about half the ingrowth.

@@

TreeMig simulates directly the mean dynamics of entire stands, but based on the  variability of light conditions within the stand   [distribution based approach, \Lischke, 1998, aggregation. To come up with the required 200 random patches of 1 ha, temporal and spatial variability intrinsic to TreeMig were combined.  Temporal variability was given by sampling from each decade of the last 100 years of the simulation, i.e. years 900-1000. For each of theses 10 decades, 20 grid cells were sampled from the simulated grid, reflecting spatial variability, results from demographic stochasticity and short-range spatial interactions

@@

In contrast to gap models, TreeMig is based on the concept of frequency distributions of tree densities classes in different tree heights on ca. 1/12 ha (833m^2) patches, and the resulting frequency distributions in discrete light. During the simulation, in each height class the frequencies of the different light classes were recorded, as well as the ingrowth of each species in the height-light class. By sampling 12 times from the height-light class ingrowths according to the height-light class frequency distribution and summing up, the stochastic ingrowths into all TreeMig height classes for one ha were calculated for each grid cell.  Then the height class boundaries were translated into DBH and the ingrowths into the height classes linearly interpolated to 7cm and 10cm. 
The ingrowth basal areas into the height classes were obtained by multiplying the interpolated ingrowth numbers with the basal areas at  7cm and 10cm, i.e. r.trees * Pi*  (0.07/2)2  and r.trees * Pi*  (0.1/2)2 .
The species specific basal area (at 1.37m) sums are a standard output of TreeMig, given by multiplying the state variables (number of trees per grid cell, height class, and species) with  the species specific basal area of each height class. 
To account for the ingrown trees dying within a decade, as an approximation the survival in each year to the power of 5 was used. Thereby, it was assumed that the trees have to survive in average 5 years in each decade, and that the current mortality is valid for these 5 years.

@@

@@

We used the process-based mechanistic model LPJ-GUESS to simulate
vegetation dynamics considering tree species, age cohorts, gap dynamics
and biogeochemical cycles. Parameters of the model reflect an updated
version of the most common European tree species as well as typical
shrub PFTs as described by Hickler et al. (2012). A detailed description
of LPJ-GUESS is available in Smith et al. 2001 and Smith et al. 2014.

@@

What climate data have you used, and how have you created your model
climate input parameters? We used the provided daily time series of
climatic variables from 1981 to 2018. The climatic variables base on the
CHELSA data set with a spatial resolution of 0.008° x 0.008° (\~ 1 km)
(Karger et al., 2021). Daily average temperature (tas) [°C], daily
average maximum temperature (tasmax) [°C], daily average minimum
temperature (tasmin) [°C], precipitation sum (pr) [mm] and surface solar
(shortwave) radiation (rsds) were taken to set up the LPJ-GUESS
simulation. Used climatic variables were converted to netCDF files. As
LPJ-GUESS requires the climatic variables in specific units, temperature
variables (tas, tasmax, tasmin) [°C] were converted to Kelvin and
surface solar radiation (rsds) [J/m2] was converted to W/m2 by dividing
rsds by 86400 s (3600 s/h \* 24 h). Nitrogen deposition was considered
to be constant at a level of 10 kgN/ha/year. Atmospheric carbon dioxide
concentration data from 1900 to 2018 was taken from the global carbon
project (Quéré et al. 2018). We used the same default LPJ-GUESS soil
code for a medium textured soil type (LPJ-GUESS soil class 2) for each
grid cell (Sitch et al. 2003).

@@

@@

@@

How have you defined the species composition in your model simulation?
(e.g., species levels, PFTs, etc.)? Please describe the species
categories used. LPJ-GUESS can be run in different modes, with different
levels of abstraction of the population and community processes. For
this study, the more detailed 'cohort' mode was used, in which
individuals, patches and vertical canopy structure are represented
explicitly, but living individuals within a cohort (age class) of a
given tree species in a given patch are assumed to be identical (in
terms of all state variables, such as height and stem diameter).
Simulations consider the main tree species as well as other PFTs
yielding an appropriate representation of European vegetation, including
only tree and shrub species that are widely distributed across Europe
and can become dominant in some areas (Hickler et al., 2012).
Mediterranean rain-green small shrubs with shallow roots (e.g. Lavendula
spp., Rosmarinus spp.) and alpine/arctic shrubs (e.g. Vaccinium spp.)
were represented as PFTs in the model. As in earlier applications of
LPJ-GUESS, herbaceous vegetation was represented by two 'generic herb'
PFTs, with C3 and C4 photosynthesis, respectively. The final set
included 16 tree species, one Mediterranean shrub PFT, one boreal/alpine
shrub PFT, and the two herbaceous PFTs described earlier (Table 2). Acer
pseudoplatanus and Alnus glutinosa are not parameterized in LPJ-GUESS
and therefore we did not simulate these species.

@@

How have you defined the simulation area for your model (e.g., size and
other relevant site aspects)? In this study, vegetation is simulated in
so called patches that represent forest areas of fixed size. Patches can
be regarded as sample stands in a grid cell, each of which represents an
0.1 ha (comparable to figure 1 of the tree regeneration workshop
protocol). In this study, vegetation dynamics were simulated by 2000
replicate patches each 0.1 ha in size. To match the required sample
size, we processed the data so that ten 0.1 ha patches were merged to
one 1 ha sample.

@@

What are the relevant aspects of the dispersal (if any) and seed input
parameters in your model? Model formulations of establishment are based
on those employed within the 'forest gap' model FORSKA (Leemans &
Prentice, 1989; Prentice et al., 1993). The number of new saplings of
each woody PFT/species and in each patch each year is drawn at random
from the Poisson distribution, with an expectation influenced by a
PFT/species-specific maximum establishment rate and by the 'propagule
pool', i.e. the amount of carbon allocated to reproduction by all
individuals of the PFT/species at all patches in the previous year. No
saplings are established in a given patch if the minimum PAR level at
the forest floor is below a PFT/species-specific threshold, which is
higher for more light-demanding PFTs/species.

Table 1: Parameter settings associated with vegetation establishment.
Parameter Description Parameter setting in this study ifbgestab Whether
background establishment enabled (0,1) 1 ifsme Whether spatial mass
effect enabled for establishment (0,1) 1 ifstochmort Whether stochastic
mortality is enabled (0,1) 1 ifstochestab Whether stochastic
establishment is enabled (0,1) 1 ifdroughtlimitedestab Whether
establishment is limited by growing season drought (0,1) 1 tcmin_est Min
20-year coldest month mean temp for establishment (deg C) PFT/species
specific (see table 2 and 3) tcmax_est Max 20-year coldest month mean
temp for establishment (deg C) PFT/species specific (see table 2 and 3)
twmin_est Min warmest month mean temp for establishment (deg C)
PFT/species specific (see table 2 and 3) twminusc Continentality
parameter for boreal summer-green trees 0 gdd5min_est Min GDD on 5 deg C
base for establishment PFT/species specific (see table 2 and 3) est_max
Max sapling establishment rate (indiv/m2/year) PFT/species specific (see
table 2 and 3) gdd0_min min GDD0 for survival/establishment reestab Pft
reestablishment ("NONE", "RESTRICTED" or "ALL") ALL

@@

What is the simulation length, and how have you determined the
simulation time until the equilibrium? The simulations were initialized
from bare ground (no biomass) and the model was spun up for 500 years
until the modelled vegetation was in approximate equilibrium with the
climate and CO2 around the year 1900 (Smith et al., 2001; Hickler et
al., 2012).

@@

### Output v

Please describe how have you calculated each of the variables in the
output data. The generated output file is called
Output_SGN_LPJGUESS.txt. Each row of the output dataset contains
variables regarding the decade 2008-2018. The first two columns describe
the study site (column "site", values reflect site ID) and the 1 ha
samples of that specific site (column "sample", values range from 1 to
200). Output variables are species-specific. Although we simulated a
total of 20 European tree species and PFTs, our output file considers
only the species required by the protocol (column "species", values are
the abbreviations for the species name ("abal" for Abies alba, "cabe"
for Carpinus betulus, "fasy" for Fagus sylvatica, "frex" for Fraxinus
excelsior, "pabi" for Picea abies, "pisy" for Pinus sylvestris , "tico"
for Tilia cordata , "betu_pen" for Betula pendula, "betu_pen" for Betula
pubescens, "quer_coc" for Quercus coccifera, "quer_il" for Quercus ilex,
"quer_pub" for Quercus pubescens, and "quer_rob" for Quercus robur).
Since Acer pseudoplatanus ("acps") and Alnus glutinosa ("algl") are not
parameterized for LPJ-GUESS so far, we have not included these species
in the output. The first variable is the number of 7 cm recruits
"r_7cm.trees" calculated as the sum of all trees of the same species
that cross the threshold of 7 cm DHB between 2008 and 2018 in that
specific sample of that specific site. The variable "r_10cm.trees"
refers to 10 cm threshold respectively. The 7 cm recruits' basal area
„r_7cm.ba" is calculated as a weighted sum of r_7cm.trees multiplied by
recruits basal area at that time (with ba=(dbh/2)∙π). Similarly,
r_10cm.ba was calculated for 10 cm recruits. The last two columns
describes the mean annual basal area and the mean number of trees per
site, sample, and species overall years between 2008 and 2018 (column
"ba").

@@

Table 2: Species/PFT characteristics and parameter of Hickler et al.
2012. r base respiration rate, kallm1 constant in allometry equations,
Tc,min minimum winter temperature for establishment, Tc,max maximum
coldest-month temperature for establishment, GDD5 minimum degree-day sum
above 5 °C for establishment, fAWC minimum growing-season fraction of
available soil water holding capacity in the first soil layer, rfire
fraction of individuals surviving fire, kla:sa leaf longevity, CAmax
maximum woody crown area. Parameter that are associated with
establishment processes are marked in bold.

 Table 3: Parameters associated with different shade tolerance classes
(updates compared to Hickler et al. 2012). Parameter that are associated
with establishment processes are marked in bold.

@@

Sitch, S., Smith, B., Prentice, I. C., Arneth, A., Bondeau, A., Cramer,
W., ... & Venevsky, S. (2003). Evaluation of ecosystem dynamics, plant
geography and terrestrial carbon cycling in the LPJ dynamic global
vegetation model. Global change biology, 9(2), 161-185.

@@

Soil quality data was converted to maximum rooting depth. Specifically, the soil quality value multiplied by 2 and the result was used as maximum rooting depth. We used the multiplication, as we typically use maximum soil depth and maximum rooting depth of 10m. Previous model simulations showed that rooting depth and soil dept have strong impacts on vegetation dynamics by influencing water availability to plants (Langan et al., 2017). We therefore considered it as adequate representation of soil quality. Other soil characteristics including volumetric water-holding capacity, soil hydraulic conductivity, soil bulk density, soil depth, soil texture, soil carbon content, soil wilting point and field capacity were taken from the FAO (http://www.fao.org, Nachtergaele et al., 2009).

@@

### Dispersal 
The aDGVM2 does not simulate seed dispersal.

@@

\newpage
\beginsupplement

#  Supplementary Material 3: Supplementary figures and tables

>>>>

---
title: "Supplementary figures\nCapturing regeneration in models of forest dynamics"
author: 
- Olalla Díaz-Yáñez, 
- Yannek Käber, 
- Mats Mahnken, 
- Samuel Fischer, 
- Xavier Morin, 
- Manfred Lexer, 
- Christian Hochauer, 
- Katarina Merganicova, 
- Tobias Mette, 
- Werner Rammer, 
- Kristin H. Braziunas, 
- Paola Mairota, 
- Josef Bruna,
- Heike Lischke, 
- Daniel Scherrer, 
- Simon Scheiter, 
- Thomas Hickler, 
- Tim Anders,
- Jessica Hetzer, 
- Harald Bugmann
date: "`r Sys.Date()`"
output: 
  pdf_document:
    fig_caption: yes
    keep_tex: yes
    number_sections: yes
    latex_engine: xelatex
    extra_dependencies: ["flafter", "float"]
    toc: yes
bibliography: Regeneration.bib
link-citations: yes
linkcolor: blue
csl: global-change-biology.csl
editor_options: 
  markdown: 
    wrap: 72
header-includes:
- \usepackage[left]{lineno}
- \linenumbers
- \newcommand{\beginsupplement}{\setcounter{table}{0}  \renewcommand{\thetable}{S\arabic{table}} \setcounter{figure}{0} \renewcommand{\thefigure}{S\arabic{figure}}}
---

@@

## 4C

@@

Morin X., Damestoy T., Toïgo M., Jactel H., Castagneyrol B. Meredieu C.
2020. Using forest gap models and experimental data to explore long-term
effects of tree diversity on the productivity of mixed planted forests.
Annals of Forest Science. 77
<https://doi.org/10.1007/s13595-020-00954-0>.

@@

Soil quality is represented by a continuous value between 1 and 5, as
provided (see 1.2.3) Soil Field Capacity (SFC) was assessed as follows:
SFC = 8+(X-1)\*(25-8)/(5-1), in which X is the 'soil_qual' value in the
provided 'soil_quality_200.csv' file.

@@

Outputs consisted in species specific regeneration rates per decade and
per ha, by considering the last 2000 years of simulations to ensure that
pseudo-equilibrium state is reached. Results were aggregated for every
10 years for the last 2000 years at the 1ha-level. Variables: - Site:
numbers corresponding to the climate data file - Sample: one per decade,
for a total of 2000 years = identified from 1 to 200 - Species: name
specified as requested (eg. 'fasy') - r.trees: number of trees
outcrossing the threshold (7 or 10 cm) during the decade, for the 10
patches simulated (= 1ha), per species - r.ba: summed basal area of the
r.trees, per species - ba: summed basal area of all trees per species at
the end of the decade - dbh: threshold used

@@

```{=html}
<!--
@article{fischer_lessons_2016,
    title = {Lessons learned from applying a forest gap model to understand ecosystem and carbon dynamics of complex tropical forests},
    volume = {326},
    issn = {03043800},
    url = {https://linkinghub.elsevier.com/retrieve/pii/S0304380015005505},
    doi = {10.1016/j.ecolmodel.2015.11.018},
    language = {en},
    urldate = {2022-09-15},
    journal = {Ecological Modelling},
    author = {Fischer, Rico and Bohn, Friedrich and Dantas de Paula, Mateus and Dislich, Claudia and Groeneveld, J{\"u}rgen and Guti{\'e}rrez, Alvaro G. and Kazmierczak, Martin and Knapp, Nikolai and Lehmann, Sebastian and Paulick, Sebastian and P{\"u}tz, Sandro and R{\"o}dig, Edna and Taubert, Franziska and K{\"o}hler, Peter and Huth, Andreas},
    month = apr,
    year = {2016},
    pages = {124--133},
    file = {Volltext:C\:\\Users\\fischsam\\Zotero\\storage\\EISNTZF6\\Fischer et al. - 2016 - Lessons learned from applying a forest gap model t.pdf:application/pdf},
}
-->
```
A full description of the model can be found in @fischer_lessons_2016.
The basis for the parameterization we applied is described in @bohn2014.

@@

To incorporate climate, we used the data set
\`daily_era5_chelsa_200.csv', which contains climate data on a daily
resolution. We considered the covariates precipitation, temperature, and
irradiance, the latter of which we converted to units used in Formind.
Adding on to these data, we used a constant and site-independent
potential evapotranspiration (PET) value of
$4\frac{\text{mm}}{\text{d}}$ and assumed that the mean day length was
$12\text{h}$ at all sites.

@@

Bugmann, H. (1994). On the ecology of mountainous forests in a changing
climate: A simulation study. Huber, N., Bugmann, H., & Lafond, V.
(2020). Capturing ecological processes in dynamic forest models: Why
there is no silver bullet to cope with complexity. Ecosphere, 11(5).
<https://doi.org/10.1002/ecs2.3109>

@@

Soil data were translated to bucket sizes values based on the
assumpation that the lowest soil quality value of 1 refers to a bucket
size of 10 cm and the highest soil quality of 5 refers to a bucket size
of 35 cm. Specifically we used the following formula to calculate the
bucket size bucket_size = 10 + (soil_quality-1) /4  \* 25.

@@

@@

Please noticed that we included 0 values for species missing in a given
sample and for samples in sites where Landis did not simulate any tree.

@@

Heike Lischke

@@

As climatic input, the monthly time series of mean temperature and
precipitation sum were used. With these climatic variables and
additional data about slope, aspect and soil field capacity, yearly time
series of the following bioclimatic variables were calculated: (1)
DDSum, the yearly day degree sum above 5.5°C , (2) MinWiT, the lowest
mean monthly temperature, (3) DrStr, drought stress according to a
bucket model that estimates a monthly PET following (Thornthwaite and
Mather 1957), using temperature, latitude, slope and aspect as input,
and calculates the monthly soil water content based on soil water
content in the previous month, precipitation, interception, and field
capacity ("bucketsize") (Bugmann 1994, Fischlin et al. 1995). Drought
stress is then given by 1- the ratio between demand (PET-interception)
and the supply, a function of the soil water content. To get climate
data also for the spin-up, bioclimate was calculated from the monthly
data as given, and then extended by sampling from the given data from
the end of the climate data to get 1000 years. The same sequence of
sampling years was used in all simulations and sites.

@@

The soil bucketsize (available water capacity of the upmost 1m soil
layer) required for the drought stress calculation, was derived from the
soil quality data set, by Bucketsize = 10 + (soil_quality-1) /4 \* 25 ,
to covert the range from 10 cm to 35cm

@@

Slope and aspect were taken from the topographic data set and converted
to a common variable entering the PET calculation, by slasp=
2*cos(aspect* pi/180) \*min(1, slope/60)

@@

TreeMig simulates the full feedback from seed production to ingrowth.
Some of the involved processes depend on environment. (1) Overall seed
production in TreeMig is described by multiplying a species specific
seed production per tree that depends on tree size and thus indirectly
on the growth, which depends on DDsum, drought stress, nutrients and
light and the adult tree number that depends on survival that depends on
DDsum, drought stress, nutrients and light. (2) Seed dispersal is
independent of environment. (3) Seed bank dynamics -- also environment
independent - is given by seed input, seed mortality, and germination.
Additionally, the number of seeds S of a species in the seed bank of
this species (SB) is limited to a carrying capacity C of 1000, by\
SB = C + (SB - C) \* Exp(-S/C).\
(4) Germination and seedling survival depend on browsing, drought,
winter temperature, DDSum and light. (5) The saplings up to 1.37 die and
grow like adult trees, depending on DDsum, drought stress and light.
Only the parameter value of the light dependence differs from that of
the adult trees.\
The regeneration processes were simulated annually in the order 1.
germination, 2. germinated seeds leave the seed bank, 3. mortality of
seeds in the seed bank, 4. seed entry by seed dispersal. The ingrowth
numbers were very sensitive to this order. Simulations with another
sequence, in which the seeds in the seed bank died first and then the
remaining ones germinated, resulted in about half the ingrowth.

@@

TreeMig simulates directly the mean dynamics of entire stands, but based
on the variability of light conditions within the stand [distribution
based approach, Lischke, 1998, aggregation. To come up with the required
200 random patches of 1 ha, temporal and spatial variability intrinsic
to TreeMig were combined. Temporal variability was given by sampling
from each decade of the last 100 years of the simulation, i.e. years
900-1000. For each of theses 10 decades, 20 grid cells were sampled from
the simulated grid, reflecting spatial variability, results from
demographic stochasticity and short-range spatial interactions

@@

In contrast to gap models, TreeMig is based on the concept of frequency
distributions of tree densities classes in different tree heights on ca.
1/12 ha (833m\^2) patches, and the resulting frequency distributions in
discrete light. During the simulation, in each height class the
frequencies of the different light classes were recorded, as well as the
ingrowth of each species in the height-light class. By sampling 12 times
from the height-light class ingrowths according to the height-light
class frequency distribution and summing up, the stochastic ingrowths
into all TreeMig height classes for one ha were calculated for each grid
cell. Then the height class boundaries were translated into DBH and the
ingrowths into the height classes linearly interpolated to 7cm and 10cm.
The ingrowth basal areas into the height classes were obtained by
multiplying the interpolated ingrowth numbers with the basal areas at
7cm and 10cm, i.e. r.trees \* Pi\* (0.07/2)2 and r.trees \* Pi\*
(0.1/2)2 . The species specific basal area (at 1.37m) sums are a
standard output of TreeMig, given by multiplying the state variables
(number of trees per grid cell, height class, and species) with the
species specific basal area of each height class. To account for the
ingrown trees dying within a decade, as an approximation the survival in
each year to the power of 5 was used. Thereby, it was assumed that the
trees have to survive in average 5 years in each decade, and that the
current mortality is valid for these 5 years.

@@

Jessica Hetzer, Thoms Hickler, Tim Anders

@@

The model was originally developed by Ben Smith of Lund University in a
collaboration also involving the Potsdam Institute for Climate Impact
Research and the Max-Planck Institute for Biogeochemistry. Over the
years, many people from institutes around the world have contributed to
the refinement and further development of the model.

We used the process-based mechanistic model LPJ-GUESS to simulate
vegetation dynamics considering tree species, age cohorts, gap dynamics
and biogeochemical cycles. Parameters of the model reflect an updated
version of the most common European tree species as well as typical
shrub PFTs as described by Hickler et al. (2012). A detailed description
of LPJ-GUESS is available in Smith et al. 2001 and Smith et al. 2014.
Source code is available on demand (see
<https://web.nateko.lu.se/lpj-guess/>).

@@

We used the provided daily time series of climatic variables from 1981
to 2018. The climatic variables base on the CHELSA data set with a
spatial resolution of 0.008° x 0.008° (\~ 1 km) (Karger et al., 2021).
Daily average temperature (tas) [°C], daily average maximum temperature
(tasmax) [°C], daily average minimum temperature (tasmin) [°C],
precipitation sum (pr) [mm] and surface solar (shortwave) radiation
(rsds) were taken to set up the LPJ-GUESS simulation. Used climatic
variables were converted to netCDF files. As LPJ-GUESS requires the
climatic variables in specific units, temperature variables (tas,
tasmax, tasmin) [°C] were converted to Kelvin and surface solar
radiation (rsds) [J/m2] was converted to W/m2 by dividing rsds by 86400
s (3600 s/h \* 24 h). Nitrogen deposition was considered to be constant
at a level of 10 kgN/ha/year. Atmospheric carbon dioxide concentration
data from 1900 to 2018 was taken from the global carbon project (Quéré
et al. 2018).

@@

All sites We used were simulated withdefault values for a medium
textured soil type (sand: 0.35, clay: 0.15, silt: 0.5) in LPJ-GUESS.

@@

Topography was not considered.

@@

LPJ-GUESS can be run in different modes, with different levels of
abstraction of the population and community processes. For this study,
the more detailed 'cohort' mode was used, in which individuals, patches
and vertical canopy structure are represented explicitly, but living
individuals within a cohort (age class) of a given tree species in a
given patch are assumed to be identical (in terms of all state
variables, such as height and stem diameter). Simulations consider the
main tree species as well as other PFTs yielding an appropriate
representation of European vegetation, including only tree and shrub
species that are widely distributed across Europe and can become
dominant in some areas (Hickler et al., 2012). Mediterranean rain-green
small shrubs with shallow roots (e.g. Lavendula spp., Rosmarinus spp.)
and alpine/arctic shrubs (e.g. Vaccinium spp.) were represented as PFTs
in the model. As in earlier applications of LPJ-GUESS, herbaceous
vegetation was represented by two 'generic herb' PFTs, with C3 and C4
photosynthesis, respectively. The final set included 16 tree species,
one Mediterranean shrub PFT, one boreal/alpine shrub PFT, and the two
herbaceous PFTs described earlier (Table 2). To increase the chances of
shade-intolerant species becoming established, patch-destroying
disturbances were allowed as described in Hickler et al. 2012. Acer
pseudoplatanus and Alnus glutinosa are not parameterized in LPJ-GUESS
and therefore not considered.we did not simulate these species.

@@

In this study, vegetation is simulated in so called patches that
represent forest areas of fixed size. Patches can be regarded as sample
stands in a grid cell, each of which represents an 0.1 ha (comparable to
figure 1 of the tree regeneration workshop protocol). In this study,
vegetation dynamics were simulated by 2000 replicate patches each 0.1 ha
in size. To match the required sample size, we processed the data so
that ten 0.1 ha patches were merged to one 1 ha sample.

@@

Model formulations of establishment are based on those employed within
the 'forest gap' model FORSKA (Leemans & Prentice, 1989; Prentice et
al., 1993). The number of new saplings of each woody PFT/species and in
each patch each year is drawn at random from the Poisson distribution,
with an expectation influenced by a PFT/species-specific maximum
establishment rate and by the 'propagule pool', i.e. the amount of
carbon allocated to reproduction by all individuals of the PFT/species
at all patches in the previous year. No saplings are established in a
given patch if the minimum PAR level at the forest floor is below a
PFT/species-specific threshold, which is higher for more light-demanding
PFTs/species.

Table 1: Parameter settings associated with vegetation establishment.

![](images/image-1265081375.png)

@@

What is the simulation length, and how have you determined the
simulation time until the equilibrium?

The simulations were initialized from bare ground (no biomass) and the
model was spun up for 500 years until the modelled vegetation was in
approximate equilibrium with the climate and CO2 around the year 1900
(Smith et al., 2001; Hickler et al., 2012).

@@

Please describe how have you calculated each of the variables in the
output data.

The generated output file is called Output_SGN_LPJGUESS.txt. Each row of
the output dataset contains variables regarding the decade 2008-2018.
The first two columns describe the study site (column "site", values
reflect site ID) and the 1 ha samples of that specific site (column
"sample", values range from 1 to 200). Output variables are
species-specific. Although we simulated a total of 20 European tree
species and PFTs, our output file considers only the species required by
the protocol (column "species", values are the abbreviations for the
species name ("abal" for Abies alba, "cabe" for Carpinus betulus, "fasy"
for Fagus sylvatica, "frex" for Fraxinus excelsior, "pabi" for Picea
abies, "pisy" for Pinus sylvestris , "tico" for Tilia cordata ,
"betu_pen" for Betula pendula, "betu_pen" for Betula pubescens,
"quer_il" for Quercus ilex, "quer_pub" for Quercus pubescens, and
"quer_rob" for Quercus robur). Since Acer pseudoplatanus ("acps") and
Alnus glutinosa ("algl") are not parameterized for LPJ-GUESS so far, we
have not included these species in the output. Further we excluded
shrubs (alpine/arctic shrubs and Quercus coccifera) and herbaceous
vegetation from the final output as this study focusses on tree
regeneration only. The first variable is the number of 7 cm recruits
"r_7cm.trees" calculated as the sum of all trees of the same species
that cross the threshold of 7 cm DHB between 2008 and 2018 in that
specific sample of that specific site. The variable "r_10cm.trees"
refers to 10 cm threshold respectively. The 7 cm recruits' basal area
„r_7cm.ba" is calculated as a weighted sum of r_7cm.trees multiplied by
recruits basal area at that time (with ba=(dbh/2)∙π). Similarly,
r_10cm.ba was calculated for 10 cm recruits. The last two columns
describes the mean annual basal area and the mean number of trees per
site, sample, and species overall years between 2008 and 2018 (column
"ba").

@@

Table 2: Species/PFT characteristics and parameter of Hickler et al.
2012. r base respiration rate, kallm1 constant in allometry equations,
Tc,min minimum winter temperature for establishment, Tc,max maximum
coldest-month temperature for establishment, GDD5 minimum degree-day sum
above 5 °C for establishment, fAWC minimum growing-season fraction of
available soil water holding capacity in the first soil layer, rfire
fraction of individuals surviving fire, kla:sa leaf longevity, CAmax
maximum woody crown area. Parameter that are associated with
establishment processes are marked in bold. Species/PFT

![](images/image-2144349832.png)

@@

@@

Soil quality data was converted to maximum rooting depth. Specifically,
the soil quality value multiplied by 2 and the result was used as
maximum rooting depth. We used the multiplication, as we typically use
maximum soil depth and maximum rooting depth of 10m. Previous model
simulations showed that rooting depth and soil dept have strong impacts
on vegetation dynamics by influencing water availability to plants
(Langan et al., 2017). We therefore considered it as adequate
representation of soil quality. Other soil characteristics including
volumetric water-holding capacity, soil hydraulic conductivity, soil
bulk density, soil depth, soil texture, soil carbon content, soil
wilting point and field capacity were taken from the FAO
(<http://www.fao.org>, Nachtergaele et al., 2009).

@@

### Dispersal

The aDGVM2 does not simulate seed dispersal.

@@

```{=tex}
\newpage
\beginsupplement
```
# Supplementary Material 3: Supplementary figures and tables
